---
title: "Week 13a"
subtitle: "Iterative coding"
author: "Dr. Nyssa Silbiger"
institute: "CSUN Spring 2021"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
      ratio: '15:10'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, htmltools.preserve.raw = FALSE)
library(anicon)
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{yellow}{rgb}{1, 0.8, 0.16078431372549}$$
$$\require{color}\definecolor{orange}{rgb}{0.96078431372549, 0.525490196078431, 0.203921568627451}$$
$$\require{color}\definecolor{green}{rgb}{0, 0.474509803921569, 0.396078431372549}$$
</div>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      yellow: ["{\\color{yellow}{#1}}", 1],
      orange: ["{\\color{orange}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>

<style>
.yellow {color: #FFCC29;}
.orange {color: #F58634;}
.green {color: #007965;}
</style>


```{r flair_color, echo=FALSE, warning=FALSE, message=FALSE}
library(flair)
yellow <- "#FFCC29"
orange <- "#F58634"
green <- "#007965"
```

---
# Outline of class

Working with iterative data  

1. for loops
2. map functions from purrr (more cat puns!)


<img src = "https://paulvanderlaken.files.wordpress.com/2018/12/pp550x5501.jpg", width=40%/>
---
# Set-up your script

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
```

### load data


---
# for loops

**for loops** are one way to run an iterative process (do something over and over again). However, with the *tidyverse* you can avoid needing loops for almost any basic coding needs.  Where it does come in handy is in population modeling and Bayesian modeling, for example. Also, is it a integral to programming and is happening "under the hood" whether you use them or not.

Also.... people be VERY opinionated on learning for loops.

[Twitter war on for loops](https://twitter.com/NSilbiger/status/1385696966445727748)


---
# for loops

For loops have two major parts:an indexing statement and a command (or set of commands) to repeat.  

The command looks like this:  
`for(index in sequence){`  
  `command to repeat}`
---
# {purrr}

[purr cheatsheet](https://github.com/rstudio/cheatsheets/blob/master/purrr.pdf)
---
# Totally awesome R package

[{gganatogram}](https://r-charts.com/miscellaneous/gganatogram/)


---
class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).  
Many slides modified from the following resources:   

[forcats](https://forcats.tidyverse.org/)  
[R4DS](https://r4ds.had.co.nz/factors.html)  
[Jenny Bryan Stat 545](https://stat545.com/factors-boss.html)  

