---
title: "Week 5a"
subtitle: "Data wrangling: lubridate dates and times"
author: "Dr. Nyssa Silbiger"
institute: "CSUN Spring 2021"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
      ratio: '15:10'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{yellow}{rgb}{1, 0.8, 0.16078431372549}$$
$$\require{color}\definecolor{orange}{rgb}{0.96078431372549, 0.525490196078431, 0.203921568627451}$$
$$\require{color}\definecolor{green}{rgb}{0, 0.474509803921569, 0.396078431372549}$$
</div>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      yellow: ["{\\color{yellow}{#1}}", 1],
      orange: ["{\\color{orange}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>

<style>
.yellow {color: #FFCC29;}
.orange {color: #F58634;}
.green {color: #007965;}
</style>


```{r flair_color, echo=FALSE}
library(flair)
yellow <- "#FFCC29"
orange <- "#F58634"
green <- "#007965"
```

---
# Outline of class

Learning about dates and times from the {lubridate} package 

.center[
<img src = "http://lubridate.tidyverse.org/logo.png", width=20%/>
]

1. Understanding date formats in base R

1. Converting and manipulating dates and times using {lubridate}

Lab 
1. Practice with dates

[Download this cheatsheet on lubridate](https://evoldyn.gitlab.io/evomics-2018/ref-sheets/R_lubridate.pdf)

---
# Review

1. What function do you use to join two dataframes together while preserving everything in the left dataframe?  

---
# Start your script.  
You know what to do..... 

```{r, eval = FALSE}
install.packages("lubridate") # package to deal with dates and times
```

```{r}
library(tidyverse)
library(here)
library(lubridate)
```

.center[
<img src = "https://media4.giphy.com/media/3o7aCTfyhYawdOXcFW/giphy.gif", width=60%/>
]

---
# What time is it now?

Even though this may seem silly if you have a clock, it is very helpful if you want to time stamp something in your code.

```{r}
now() #what time is it now?
```

--

 You can also ask the time in other time zones.
 
```{r}
now(tzone = "EST") # what time is it on the east coast
now(tzone = "GMT") # what time in GMT
```

---
# What time is it now?

If you only want the date and not the time

```{r}
today()
```

```{r}
today(tzone = "GMT")
```

--

 You can also ask if it is morning or night right now and if it is a leap year.
 
```{r}
am(now()) # is it morning?

leap_year(now()) # is it a leap year?
```

---
# Date specifications for lubridate

{lubridate} does a good job of guessing the format of your date if you give just a little hint.

First, your dates **must be a character**.  

A common mistake is that you get an error trying to convert a date when it is a factor. Always check to make sure it is a character (i.e. put quotes around it).



| Date| function   |       
|----------|:-------------:|
| 2021-02-24 | ymd()|
| 02/24/2021 | mdy()|
| February 24 2021 | mdy()|
| 24/02/2021 | dmy()|


y = year  
m = month  
d = day  

---
# Date specifications for lubridate

These will all produce the same results as ISO dates 
```{r}
ymd("2021-02-24")
mdy("02/24/2021")
mdy("February 24 2021")
dmy("24/02/2021")

```

---
# Date specifications for lubridate

Say you wanted to label something as just the month, day, or year (not a full date).

```{r}
#create a month
month(2)

# save it as feb
month(2, label = TRUE) 

# spell out the month
month(2, label = TRUE, abbr = FALSE)

```
---
# Date and Time specifications with lubridate

| Time| function   |       
|----------|:-------------:|
| 2021-02-24 10:22:20 PM | ymd_hms()|
|02/24/2021 22:22:20 | mdy_hms()|
|February 24 2021 10:22 PM | mdy_hm()|


y = hour   
m = minute (1st m)  
d = second  
h = hour  
m = minute (2nd m)  
s = second 

---

# Date and Time specifications with lubridate

```{r}
ymd_hms("2021-02-24 10:22:20 PM")

mdy_hms("02/24/2021 22:22:20")

mdy_hm("February 24 2021 10:22 PM")

```
---

---
Lab

Practice joins on your own.  You do not have to turn anything in for this lab. 

Reminder: We are starting RMarkdown next week. If you have not yet completed the markdown lab from week 2, do it now. I will expect that you know the basics from that lab assignment **before** class on Monday.  


Who wants to share their Tidy Tuesday from last week before we end?

---
class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

Some slides modified from [Tidy Explain ](https://github.com/gadenbuie/tidyexplain#:~:text=A%20mutating%20join%20allows%20you,one%20table%20to%20the%20other.)


