---
title: "Week 4b"
subtitle: "Data wrangling: tidyr"
author: "Dr. Nyssa Silbiger"
institute: "CSUN Spring 2021"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
      ratio: '15:10'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{yellow}{rgb}{1, 0.8, 0.16078431372549}$$
$$\require{color}\definecolor{orange}{rgb}{0.96078431372549, 0.525490196078431, 0.203921568627451}$$
$$\require{color}\definecolor{green}{rgb}{0, 0.474509803921569, 0.396078431372549}$$
</div>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      yellow: ["{\\color{yellow}{#1}}", 1],
      orange: ["{\\color{orange}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>

<style>
.yellow {color: #FFCC29;}
.orange {color: #F58634;}
.green {color: #007965;}
</style>


```{r flair_color, echo=FALSE}
library(flair)
yellow <- "#FFCC29"
orange <- "#F58634"
green <- "#007965"
```

---
# Outline of class

1. Intro to {tidyr}
 - pivot data between long and wide formats
 - Separate and unite columns

Lab 
1. Practice with tidyr



---
# Review

1. What function do you use to subset rows by some criterion?  

1. How do I add a new column to my dataframe?
---
# Intro to the {tidyr} package (part of the TidyVerse)

.center[
<img src = "https://robotwealth.com/wp-content/uploads/2020/05/tidyr.png", width=50% />
]

---
# New dataset!

We are going to use a subset of data from Silbiger et al. 2020 Proceedings of the Royal Society: B.

In the data folder for this week in Spring-2021, you will see two .csv files:

1. chemicaldata_manualua.csv
1. chem_data_dictionary.csv

Move both of them to the data folder of YOUR repo.

Explore the data dictionary to understand what you will be working with.

Basics: I collected biogeochemistry data at sites in Hawai'i with submarine groundwater discharge (fresh groundwater that flows into coastal oceans through cracks in the reef plate). We have data on nutrient and carbonate chemistry from two sites, two seasons, high and low tide, and during the day and night.

---
# Set up your script for today
```{r, message=FALSE, warning=FALSE}
### Today we are going to practice tidy with biogeochemistry data from Hawaii ####
### Created by: Dr. Nyssa Silbiger #############
### Updated on: 2021-02-17 ####################


#### Load Libraries ######
library(tidyverse)
library(here)


### Load data ######
ChemData<-read_csv(here("Week_4","data", "chemicaldata_maunalua.csv"))
View(ChemData)
glimpse(ChemData)

```
---
# Another way to remove all the NAs

There are several ways we can remove all the NAs in a data set

```{r}
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) #filters out everything that is not a complete row

View(ChemData_clean)  

```

---
class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

Some slides modified from [Andrew Wheiss ](https://evalsp21.classes.andrewheiss.com/projects/01_lab/slides/01_lab.html#73)


