---
title: "Week 4b"
subtitle: "Data wrangling: tidyr"
author: "Dr. Nyssa Silbiger"
institute: "CSUN Spring 2021"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
      beforeInit: "macros.js"
      ratio: '15:10'
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
<div style = "position:fixed; visibility: hidden">
$$\require{color}\definecolor{yellow}{rgb}{1, 0.8, 0.16078431372549}$$
$$\require{color}\definecolor{orange}{rgb}{0.96078431372549, 0.525490196078431, 0.203921568627451}$$
$$\require{color}\definecolor{green}{rgb}{0, 0.474509803921569, 0.396078431372549}$$
</div>

<script type="text/x-mathjax-config">
MathJax.Hub.Config({
  TeX: {
    Macros: {
      yellow: ["{\\color{yellow}{#1}}", 1],
      orange: ["{\\color{orange}{#1}}", 1],
      green: ["{\\color{green}{#1}}", 1]
    },
    loader: {load: ['[tex]/color']},
    tex: {packages: {'[+]': ['color']}}
  }
});
</script>

<style>
.yellow {color: #FFCC29;}
.orange {color: #F58634;}
.green {color: #007965;}
</style>


```{r flair_color, echo=FALSE}
library(flair)
yellow <- "#FFCC29"
orange <- "#F58634"
green <- "#007965"
```

---
# Outline of class

1. Intro to {tidyr}
  - Separate and unite columns
  - Pivot data between long and wide formats

Lab 
1. Practice with tidyr



---
# Review

1. What function do you use to subset rows by some criterion?  

1. How do I add a new column to my dataframe?
---
# Intro to the {tidyr} package (part of the TidyVerse)

.center[
<img src = "https://robotwealth.com/wp-content/uploads/2020/05/tidyr.png", width=50% />
]

---
# New dataset!

We are going to use a subset of data from Silbiger et al. 2020 Proceedings of the Royal Society: B.

In the data folder for this week in Spring-2021, you will see two .csv files:

1. chemicaldata_manualua.csv
1. chem_data_dictionary.csv

Move both of them to the data folder of YOUR repo.

Explore the data dictionary to understand what you will be working with.

Basics: I collected biogeochemistry data at sites in Hawai'i with submarine groundwater discharge (fresh groundwater that flows into coastal oceans through cracks in the reef plate). We have data on nutrient and carbonate chemistry from two sites, two seasons, high and low tide, and during the day and night.

---
# Set up your script for today
```{r, message=FALSE, warning=FALSE}
### Today we are going to practice tidy with biogeochemistry data from Hawaii ####
### Created by: Dr. Nyssa Silbiger #############
### Updated on: 2021-02-17 ####################


#### Load Libraries ######
library(tidyverse)
library(here)


### Load data ######
ChemData<-read_csv(here("Week_4","data", "chemicaldata_maunalua.csv"))
View(ChemData)
glimpse(ChemData)

```
---
# Another way to remove all the NAs

There are several ways we can remove all the NAs in a data set

```{r}
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) #filters out everything that is not a complete row

View(ChemData_clean)  

```

---
# Separate function

We notice a silly column named Tide_time, which has tide (low/high) paired with time of day (day/night).  

Nyssa did NOT follow directions for tidy data from the first week of class... Now we have to fix it.

.center[
<img src ="https://media.tenor.com/images/6a3223dda90636415530c9521d4a5a2a/tenor.gif"/>
]

---
# Separate function

Notice that I have two bits of information: Low_Day is low tide during the day.  It would be way easier to plot and analyze the data if we separate those into separate columns.

hit ?separate to see all the things you can do, but the simplest use is:  

separate(  
  data = [data frame you are using],  
  col = [column that you want to separate],  
  into = [name of the new columns],  
  sep = [what are you separating by?])  

---
# Separate

Let's continue to build off our data frame
```{r}
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col #<<
           into = c("Tide","Time"), # separate it into two columns Tide and time #<<
           sep = "_" ) # separate by _ #<<

head(ChemData_clean)  

```

---
# Separate

Notice this deletes the original column.  If we wanted to keep it we would add *remove = FALSE*.

---
# Separate

Let's continue to build off our data frame
```{r}
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) # keep the original tide_time column #<<

head(ChemData_clean)  

```

Notice this deletes the original column.  If we wanted to keep it we would add *remove = FALSE*.

---
# Unite
What if we have the opposite problem and what to *unite* two columns into one? 

Let's say we want a new column that *unites* site with zone.  We use the unite() function which has the same arguments as separate() 

--

```{r}
ChemData_clean<-ChemData %>%
  filter(complete.cases(.)) %>% #filters out everything that is not a complete row
  separate(col = Tide_time, # choose the tide time col
           into = c("Tide","Time"), # separate it into two columns Tide and time
           sep = "_", # separate by _
           remove = FALSE) %>% # keep the original tide_time column
  unite(col = "Site_Zone", # the name of the NEW col #<<
         c(Site,Zone), # the columns to unite #<<
        sep = ".", # lets put a . in the middle #<<
        remove = FALSE) # keep the original #<<

head(ChemData_clean)
```


---
# Today's totally awesome R package

### Put a Bernie on it


```{r, eval=FALSE, message = FALSE}
#devtools::install_github("R-CoderDotCom/ggbernie@main")

```

```{r, message=FALSE, warning=FALSE, fig.height=4, fig.width=5}
library(ggbernie)

ggplot(ChemData_clean) +
  geom_bernie(aes(x = pH, y = NN), bernie = "sitting")
```



---
class: center, middle

# Thanks!

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

Some slides modified from [Andrew Wheiss ](https://evalsp21.classes.andrewheiss.com/projects/01_lab/slides/01_lab.html#73)


